\section*{Abstract}

The rapid development of precision livestock farming has created a critical need for integrated systems capable of performing multiple concurrent health assessments. Traditional cattle monitoring research suffers from two primary limitations: most studies address individual tasks in isolation, and they rely on private datasets that prevent reproduction and benchmarking. This thesis addresses these gaps by proposing and implementing a unified, lightweight, and spatiotemporal multi-task deep learning framework that simultaneously processes four critical monitoring tasks from video inputs: Body Condition Scoring (BCS), behavior recognition, lameness detection, and individual animal identification. Built exclusively on publicly available datasets (ScienceDB, Dryad, MmCows, CattleLameness, OpenCows2020), our model employs a shared EfficientNet-B0 backbone encoder enhanced with a Convolutional Block Attention Module (CBAM) for spatial and channel focus. We employ task-specific heads and optimized loss functions, including Consistent Rank Logits (CORAL) loss for BCS ordinal regression, Focal Loss to address severe behavior class imbalances, and Long Short-Term Memory (LSTM) sequence-tracking layers for locomotion analysis. Single-task baselines achieved strong results: a BCS Mean Absolute Error (MAE) of 0.5566 on ScienceDB, a behavior recognition Macro F1-score of 0.7445, a lameness detection AUC of 0.9829, and 86.49\% top-1 accuracy on individual cow identification. When integrated into the unified spatiotemporal multi-task model, lameness detection achieved an uncalibrated 80.00\% accuracy (which rose to 100.00\% only under post-hoc threshold tuning, serving as a preliminary proof-of-concept) and cow identification rose to 97.58\% through temporal sequence modeling and majority voting. As expected in multi-task optimization with hard parameter sharing, BCS and behavior experienced severe negative transfer (MAE 0.7827, F1 0.4948). Consequently, the current joint model serves as a foundational baseline for future gradient-balancing research rather than a system ready for immediate clinical deployment. To evaluate generalization, cross-dataset validation was conducted on the CBVD-5 dataset, revealing critical domain shift challenges. The framework reduces edge VRAM requirements by approximately 75\% compared to deploying four independent models, establishing the first complete, reproducible multi-task pipeline for cattle health monitoring using exclusively public datasets.

\vspace{1cm}
\noindent \textbf{Keywords:} Multi-Task Learning, Spatiotemporal Modeling, Cattle Health Monitoring, Body Condition Scoring, Behavior Recognition, Lameness Detection, Edge Deployment
\pagebreak