\section*{Abstract}

The fast development of precision livestock farming has created a need for systems capable of doing multiple health assessments. Current cattle monitoring research has a major problem: most studies only tackle one task at a time and use private data that other researchers can't access. This makes it nearly impossible to verify results or deploy these systems on real farms. This thesis takes a different approach. We developed a single deep learning system that handles four cattle monitoring tasks together: Body Condition Scoring, behavior recognition, lameness detection, and identifying individual animals. What sets this work apart is our commitment to transparency—we only use publicly available datasets, so anyone can reproduce and build on our work. Our approach uses a lightweight neural network (EfficientNet-B0 or MobileNetV3-Small) as a shared foundation. We train each task sequentially, which lets us combine different types of data without needing labels for every task in every image. To boost performance, we incorporated advanced techniques like CORAL (which helps with ordinal rankings) and CBAM (which helps the model focus on important features). While the system prioritizes RGB modalities, it also uses the contributions of depth information when available. This work provides the first complete, reproducible system that can run on edge devices to monitor four cattle health tasks in real-time across different farm environments.

\vspace{1cm}
\noindent \textbf{Keywords: Multi-Task Learning, Cattle Health Monitoring, Body Condition Scoring, Behavior Recognition, Lameness Detection, Deep Learning, Transfer Learning, Public Datasets, Edge Deployment}
\pagebreak