\section{Background}

The global livestock industry plays a vital role in ensuring food security, with cattle representing one of the most economically critical livestock species worldwide. According to the Food and Agriculture Organization (FAO), the global cattle population exceeds one billion heads, contributing substantially to meat and dairy production \cite{fao2018livestock}. Consequently, the sustainability, economic viability, and productivity of farming operations are directly influenced by the health and welfare of these animals. Early detection of prevalent health issues---such as poor body condition, lameness, and abnormal behavior---can mitigate global economic losses estimated at billions of dollars annually while significantly improving animal welfare standards \cite{weary2009understanding}.

Cattle nutritional status and systemic health are fundamentally reflected in the Body Condition Score (BCS). BCS is traditionally assessed through visual and tactile evaluation of fat deposits over key anatomical regions, typically using a 5-point scale with 0.25-point increments \cite{edmonson1989bcs}. While optimal body condition is directly linked to improved milk yield, reproductive efficiency, and disease resistance, manual BCS assessment is highly time-consuming, subjective, and labor-intensive, making frequent herd-wide monitoring impractical for large-scale operations. Similarly, lameness represents another critical health concern, affecting up to 25\% of dairy cattle in intensive farming systems \cite{whay2003lameness}. Lameness causes severe pain, reduces feed intake, lowers milk production, and hinders reproduction; however, timely intervention remains possible through the early detection of gait abnormalities and abnormal body movements.

Automated behavior monitoring and individual identification provide crucial insights into cattle health, stress levels, and overall welfare. Deviations in feeding patterns, rumination time, lying duration, and social interactions frequently precede the clinical manifestations of disease \cite{weary2009understanding}. Continuous 24/7 monitoring through automated computer vision systems offers a non-invasive, objective alternative to manual observation. Furthermore, individual identification is fundamental to precision livestock farming, enabling the longitudinal tracking of health records, feeding habits, and productivity metrics for each animal. Traditional identification methods, such as physical ear tags and Radio Frequency Identification (RFID) systems, require physical contact and are prone to loss or damage. Non-invasive, vision-based identification systems overcome these limitations and can be seamlessly integrated with automated health monitoring frameworks.

The convergence of computer vision, deep learning, and Internet of Things (IoT) technologies has created unprecedented opportunities for automated, real-time cattle health monitoring \cite{neethirajan2020digitalization}. Convolutional Neural Networks (CNNs) have demonstrated extraordinary success in image classification and object detection tasks, achieving human-level performance across diverse visual recognition challenges. Transfer learning further enables the application of deep learning models pre-trained on large-scale datasets, such as ImageNet, to agricultural domains where labeled data is scarce. Despite these technological advances, existing cattle monitoring systems predominantly address individual tasks in isolation. A systematic review of 33 primary research papers conducted for this thesis reveals that while multi-task learning is frequently discussed, end-to-end implementations combining multiple health monitoring tasks are virtually non-existent. This fragmentation limits the practical utility of automated monitoring systems and fails to exploit the computational and representation-learning synergies between related tasks.


\section{Rationale of the Study and Motivation}

The motivation for this research stems from critical gaps in the current state of automated cattle health monitoring systems and the escalating demand for precision livestock farming. First, a systematic analysis of the literature indicates that a substantial majority (29 of 33, $\approx 88\%$) of current studies rely on private, proprietary datasets, which presents a major barrier to reproducibility and comparative validation. Second, the majority of existing research addresses individual tasks---such as BCS, lameness detection, behavior recognition, or individual identification---in isolation, neglecting the potential benefits of Multi-Task Learning (MTL). By sharing feature representations across related tasks, MTL can significantly reduce computational overhead while potentially improving performance. This is particularly relevant as these health indicators are biologically and behaviorally interconnected; for instance, abnormal locomotion or pain from lameness directly contributes to reduced feeding time, which subsequently leads to a decline in body condition.

Furthermore, many published solutions require expensive specialized hardware, such as thermal cameras, wearable biotelemetry sensors, or depth cameras, which are often cost-prohibitive or impractical for widespread commercial deployment. This highlights a clear need for frameworks that rely primarily on standard, low-cost RGB cameras while retaining the flexibility to integrate depth information when available. Additionally, modern agricultural operations require real-time, on-site processing because rural network connectivity is frequently unreliable and cloud-based inference introduces latency. While edge deployment is recognized as an optimal solution, most existing models rely on computationally heavy architectures unsuitable for resource-constrained edge devices. Finally, there is a distinct lack of self-supervised learning applications in cattle monitoring. Most commercial farms operate 24/7 surveillance cameras that accumulate vast quantities of video footage daily. Although this abundant unlabeled footage remains unutilized, it could serve as a valuable resource for pre-training models to learn robust representations, which is particularly beneficial when labeled data is scarce. Collectively, these challenges---reproducibility issues, task fragmentation, hardware constraints, deployment limitations, and underutilized video resources---highlight the necessity of a unified, lightweight framework for multi-task cattle monitoring.


\section{Problem Statement}

Despite significant research efforts in automated cattle health monitoring, several fundamental challenges remain unresolved. Consequently, the practical deployment and real-world impact of computer vision and deep learning solutions on commercial farms remain limited.

A primary limitation is that existing systems address monitoring tasks independently, necessitating the deployment of multiple separate models to conduct comprehensive health assessments. This disjointed approach increases computational overhead, fails to exploit shared feature representations that could mutually benefit related tasks, complicates system integration and deployment, and overlooks the correlations between different health indicators (such as the relationship between poor body condition and reduced feeding activity). A systematic review of the literature confirms the absence of a unified framework that concurrently addresses body condition scoring, behavior recognition, lameness detection, and individual identification. Multi-task learning is frequently discussed conceptually but rarely implemented in end-to-end architectures.

In order to validate experimental results, compare methodologies objectively, and build upon prior literature, researchers must evaluate models on standardized datasets. However, a significant portion of cattle monitoring studies utilize proprietary or private datasets. This lack of data sharing hinders reproducibility, prevents fair performance comparisons, and restricts cumulative progress in the field. This stands in sharp contrast to general computer vision research, which benefits from standardized, publicly accessible benchmarks such as ImageNet and COCO that facilitate consistent and transparent evaluations.

Furthermore, most studies evaluate their models using data collected from a single farm or a single breed of cattle, raising questions regarding the generalizability of the proposed methods. Models evaluated under such constrained conditions risk overfitting to specific farm layouts, camera configurations, and illumination conditions, and they may fail to generalize across different breeds. Seasonal, environmental, and temporal variations are also rarely accounted for in evaluations. Crucially, cross-dataset evaluation---where a model is trained on one dataset and tested on another from an entirely different source---is seldom performed, despite being a prerequisite for demonstrating real-world viability.

Finally, while depth and thermal imaging have attracted significant interest for cattle monitoring, multimodal research is restricted by the scarcity of publicly available datasets. There are no public cattle thermal datasets, only the MmCows dataset provides synchronized Inertial Measurement Unit (IMU) data, and only two public datasets (MmCows and Dryad) offer depth information. This data scarcity forces researchers to either collect private datasets or forego multimodal approaches entirely.


\section{Objectives}

The primary objective of this research is to develop a lightweight, edge-deployable, and openly evaluated multi-task deep learning framework for automated cattle health and behavior monitoring. By leveraging public datasets and shared feature representations, the framework aims to concurrently perform multiple health-related assessments while establishing a standardized baseline.

To achieve this goal, the research sets out the following specific objectives:

\begin{enumerate}
    \item \textbf{To design a unified, lightweight multi-task architecture} with a shared backbone and specialized task-specific heads, and to evaluate sequential training strategies to optimize performance across Body Condition Scoring, behavior recognition, lameness detection, and individual identification.
    
    \item \textbf{To improve reproducibility} by conducting all evaluations exclusively on publicly available datasets and documenting preprocessing, task-specific splits, model settings, and evaluation protocols.
    
    \item \textbf{To conduct preliminary cross-dataset evaluation} to identify domain shift challenges across different farms and environmental conditions, establishing a baseline for comprehensive generalization studies in future work.
\end{enumerate}

These objectives address the reproducibility gap in agricultural AI research and create a scalable foundation for real-time cattle health monitoring across various farm environments.


\section{Methodology in Brief}

The proposed methodology follows a structured, multi-phase process to design, implement, and validate the multi-task deep learning framework. This research workflow is divided into five key phases: data preparation, architecture design, sequential task training, evaluation, and ablation studies.


\subsection{Evaluation Protocol}

Evaluation metrics are carefully tailored to each task to address their distinct outputs and operational risks. For body condition scoring, the model's performance is measured using Mean Absolute Error (MAE) alongside classification accuracy within $\pm0.25$ and $\pm0.5$ tolerances. For behavior recognition, evaluation metrics include per-class accuracy, macro-averaged F1-score, and confusion matrix analysis. For lameness detection, evaluation metrics include accuracy, F1-score, and AUC-ROC. Individual identification is assessed using Top-1 accuracy. Finally, cross-dataset evaluation is conducted by training on a primary dataset and testing on an independent dataset to rigorously assess the model's generalization capabilities.


\subsection{Ablation Studies}

A series of structured ablation experiments will be conducted to evaluate the contributions of key framework components. For backbone selection, EfficientNet-B0 will be compared against MobileNetV3-Small and ResNet-50 to determine the optimal trade-off between computational efficiency and accuracy. For the body condition scoring task, the performance of the Rank Consistent Ordinal Regression (CORAL) loss will be compared against standard Cross-Entropy loss. Regarding training strategies, experiments will compare sequential learning and joint multi-task training. The impact of multi-modal data will be discussed through a cross-dataset modality observation. Finally, the efficacy of attention mechanisms will be evaluated by comparing architectures with and without the Convolutional Block Attention Module (CBAM).


\section{Scope and Challenges}

\subsection{Scope of the Research}

To ensure feasibility and depth of investigation, this research is defined by the following boundaries:

\begin{enumerate}
    \item \textbf{Four Core Tasks:} The scope is restricted to Body Condition Scoring, behavior recognition, lameness detection, and individual identification, selected due to their practical significance and public data availability.
    
    \item \textbf{Public Datasets Only:} All experimental evaluations are conducted exclusively on publicly available datasets to ensure a more transparent and reproducibility-oriented framework, using public datasets and documented preprocessing/splits.
    
    \item \textbf{RGB as Primary Modality:} RGB imagery serves as the primary input modality, with depth information evaluated solely as an auxiliary modality for the body condition scoring task using the Dryad dataset.
    
    \item \textbf{Dairy Cattle Focus:} The study focuses primarily on Holstein dairy cattle, which represent the dominant breed in public cattle monitoring datasets.
    
    \item \textbf{Lightweight Architectures:} To ensure suitability for resource-constrained edge deployment, the backbone selection is restricted to architectures with fewer than 10 million parameters.
    
    \item \textbf{Sequential Task Training:} Frozen-backbone sequential training serves as the primary optimization strategy, while joint multi-task fine-tuning and loss balancing are explored as secondary extensions.
\end{enumerate}


\subsection{Challenges}

\textbf{Camera Geometry Conflicts}

The four tasks in this framework require fundamentally different camera viewpoints: BCS scoring performs best from a rear-view angle (capturing hip and spine morphology), while Lameness detection requires a side-view angle (to observe gait stride length and limb symmetry). Running both tasks from a single fixed camera creates an unavoidable geometric trade-off. To address this, the deployment model assumes a multi-camera setup where a single edge device loads the shared backbone once, but selectively activates the rear-view or side-view task heads based on the incoming camera feed.



\textbf{Dataset Heterogeneity}

Public cattle datasets exhibit high heterogeneity, including variations in image resolution, quality, camera placement (top-view, rear-view, and side-view), annotation protocols, breed characteristics, and farm environments. To mitigate these discrepancies, the methodology incorporates a standardized preprocessing pipeline, robust data augmentation, and domain adaptation techniques.


\textbf{Class Imbalance}

Cattle monitoring datasets frequently exhibit severe class imbalances. In lameness detection, lame individuals are heavily outnumbered by healthy cattle; similarly, extreme body condition scores (under 2.5 or over 3.5) are significantly rarer than moderate scores, and specific behaviors (such as drinking or licking) occur infrequently. To address these imbalances, the framework employs Focal Loss for classification tasks, class-weighted sampling during training, and balanced evaluation metrics.


\textbf{Temporal Modeling for Video Tasks}

While behavior recognition and lameness detection benefit from temporal context, video-based modeling increases computational overhead and introduces architectural complexity (e.g., through LSTMs or I3D networks), alongside potential discrepancies between frame-level and clip-level annotations. The framework addresses this by first utilizing frame-level classification combined with temporal aggregation, treating recurrent model integration as a subsequent extension.


\textbf{Preventing Negative Transfer}

Multi-task learning designs are susceptible to negative transfer, where optimization for one task degrades the performance of another. This typically occurs due to conflicting task gradients during joint optimization, varying optimal feature representations across tasks, and disparity in task difficulties. The primary mitigation strategy involves sequential training with a frozen backbone to decouple task optimization, supplemented by gradient normalization and dynamic loss weighting during joint fine-tuning.


\textbf{Limited Labeled Data}

Despite utilizing multiple public datasets, the total volume of labeled data remains small compared to standard computer vision benchmarks. For example, the ScienceDB BCS dataset contains 53,566 images, in contrast to the millions in ImageNet. The MmCows dataset tracks only 16 individuals over a two-week period, and the CattleLameness dataset contains a relatively small number of lame examples. To mitigate data limitations, the methodology leverages transfer learning from ImageNet, robust regularization (such as dropout and weight decay), and extensive data augmentation, with self-supervised pre-training on unlabeled video frames designated as an optional extension.
