\section{Summary of Findings}
This research successfully developed, trained, and evaluated a unified, lightweight, and spatiotemporal multi-task deep learning framework for real-time cattle health and behavior monitoring. Using EfficientNet-B0 as a shared encoder backbone combined with a CBAM attention module, the framework consolidates four distinct computer vision tasks: Body Condition Scoring (BCS), Behavior Recognition, Lameness Detection, and Cow Identification. 

The experimental results confirmed that sequence modeling with LSTM layers is highly effective for video-based tasks, achieving an uncalibrated 80.00\% accuracy on lameness detection sequences (which rose to 100.00\% accuracy and 1.0000 AUC only under post-hoc threshold tuning, serving as a preliminary proof-of-concept). For behavior recognition, the spatiotemporal setup raised the Macro F1-score from 0.3771 (frame-level multi-task) to 0.4948, although this represents a degradation compared to the single-task baseline of 0.7445 due to expected negative transfer. The multi-task model exhibits outstanding computational efficiency, reducing VRAM requirements by approximately 75\% compared to deploying four independent single-task models, making it well-suited for resource-constrained offline farm surveillance. However, due to the severe negative transfer observed in the spatial tasks, the current joint framework is presented as a foundational baseline for future gradient-balancing research rather than a system ready for immediate clinical deployment.

\section{Contributions to the Field}
This thesis provides several key contributions to the domain of Precision Livestock Farming (PLF) and agricultural computer vision:
\begin{enumerate}
    \item \textbf{A Unified Multi-Task Framework:} To the best of our knowledge, this is the first end-to-end multi-task deep learning framework that simultaneously monitors BCS, behavior, lameness, and individual cow identity from video sequences using a single shared encoder.
    \item \textbf{Commitment to Full Reproducibility:} Unlike the majority of agricultural computer vision literature which relies on private datasets, our study was designed, trained, and validated exclusively on publicly available datasets (ScienceDB, Dryad, MmCows, CattleLameness, OpenCows2020, CBVD-5), establishing a standardized benchmark for future research.
    \item \textbf{Spatiotemporal Integration with Attention:} We successfully integrated a CBAM attention layer and LSTM sequence-tracking to process both spatial anatomical shapes (spine, hip bones) and temporal locomotion strides, bridging static 2D image analysis and recurrent 3D time-series.
    \item \textbf{Edge Optimization:} The model structure was optimized to remain under 10 million parameters, enabling cost-effective, non-invasive surveillance directly on camera nodes, reducing dependency on expensive wearable sensors.
\end{enumerate}

\section{Recommendations for Future Work}
While the framework has shown strong feasibility, several areas represent crucial directions for future expansion:
\begin{itemize}
    \item \textbf{Multi-Task Gradient Conflict Optimization (GradNorm/PCGrad):} Although joint fine-tuning was successfully executed in Phase 6 by unfreezing the shared backbone, this joint optimization introduced severe task interference (negative transfer) due to conflicting gradients, which degraded performance on spatial tasks (BCS and Behavior). Future work should integrate advanced multi-task loss balancing and gradient alignment algorithms, such as GradNorm or PCGrad (Projecting Conflicting Gradients), during joint training. These methods dynamically scale or project conflicting task gradients during backpropagation to prevent destructive interference and recover the performance of single-task baselines.
    \item \textbf{Contrast Limited Adaptive Histogram Equalization (CLAHE):} Farm environments suffer from extreme outdoor lighting variances (glare, shadows, night lighting). Integrating CLAHE as a preprocessing step will enhance local contrast, standardizing raw pixels before backbone feature extraction.
    % Behavior Sequential Modeling was implemented and is no longer future work
    \item \textbf{Federated and Collaborative Learning:} To improve cross-dataset generalization (which degraded to a Macro F1-score of 0.1245 on CBVD-5), a federated learning framework should be explored. This allows models to be trained across multiple farms in a decentralized manner, learning from diverse breeds and lighting environments without centralizing sensitive proprietary data.
    \item \textbf{Synthetic Data Generation via 3D Simulation:} To overcome the severe data scarcity and temporal overfitting associated with the small 50-video lameness dataset, future research should explore building rigged 3D biomechanical cow models in physics engines (e.g., Unreal Engine). This would allow for the programmatic generation of massive, perfectly-labeled synthetic video datasets featuring diverse grades of lameness, eliminating the bottleneck of manual data collection and annotation.
\end{itemize}
